---
title: "ISLR 7"
output: html_notebook
---

```{r}
library(ISLR2)
attach(Wage)
```

Coefficients for a 4th degree polynomial:

```{r}
fit <- lm(wage ~ poly(age, 4), data = Wage)
coef(summary(fit))
```
`poly(age, 4)` is a matrix whose columns are linear combinations of `age, age^2, ...`

Found this for the mathematical definition of poly:
> A quick answer is that poly of a vector is x essentially equivalent to the QR decomposition of the matrix whose columns are powers of x 

Some more details [here](https://www.sydney.edu.au/science/chemistry/~mjtj/CHEM3117/Resources/poly_etc.pdf)

```{r}
m <- poly(age, 4)
dim(m)
dim(Wage)
```

```{r}
fit2 <- lm(wage ~ poly(age, 4, raw=TRUE), data = Wage)
coef(summary(fit2))
```

Similarly

```{r}
fit2a <- lm(wage ~ age + I(age^2) + I(age^3) + I(age^4), data = Wage)
coef(fit2a)
```
Slightly different way:

```{r}
fit2b <- lm(wage ~ cbind(age, age^2, age^3, age^4), data = Wage)
coef(fit2b)
```

Prediction:

```{r}
agelims <- range(age)
age.grid <- seq(from = agelims[1], to = agelims[2])
preds <- predict(fit, newdata = list(age = age.grid), se = TRUE)
se.bands <- cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)
```

?
```{r}
par(mfrow = c(1, 2), mar = c(4.5, 4.5, 1, 1), oma = c(0, 0, 4, 0))
plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey") > title("Degree-4 Polynomial", outer = T)
lines(age.grid, preds$fit, lwd = 2, col = "blue")
matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
```

?
```{r}
preds2 <- predict(fit2, newdata = list(age = age.grid), se = TRUE)
max(abs(preds$fit - preds2$fit))
```

ANOVA
```{r}
fit.1 <- lm(wage ~ age, data = Wage)
fit.2 <- lm(wage ~ poly(age, 2), data = Wage)
fit.3 <- lm(wage ~ poly(age, 3), data = Wage)
fit.4 <- lm(wage ~ poly(age, 4), data = Wage)
fit.5 <- lm(wage ~ poly(age, 5), data = Wage)
anova(fit.1, fit.2, fit.3, fit.4, fit.5)
coef(summary(fit.5))
```

Adding `education`:
```{r}
fit.1 <- lm(wage ~ education + age, data = Wage)
fit.2 <- lm(wage ~ education + poly(age, 2), data = Wage)
fit.3 <- lm(wage ~ education + poly(age, 3), data = Wage)
anova(fit.1, fit.2, fit.3)
```

Using `glm`:
```{r}
fit <- glm(I(wage > 250) ~ poly(age, 4), data = Wage, family = binomial)
```

?
```{r}
preds <- predict(fit, newdata = list(age = age.grid), se = T)
summary(preds)
```

?
```{r}
pfit <- exp(preds$fit) / (1 + exp(preds$fit))
se.bands.logit <- cbind(preds$fit + 2 * preds$se.fit,
preds$fit - 2 * preds$se.fit)
se.bands <- exp(se.bands.logit) / (1 + exp(se.bands.logit))
# same as the below?
preds <- predict(fit, newdata = list(age = age.grid), type = "response", se = T)
```

?
```{r}
plot(age, I(wage > 250), xlim = agelims, type = "n", ylim = c(0, .2))
points(jitter(age), I((wage > 250) / 5), cex = .5, pch = "|", col = "darkgrey")
lines(age.grid, pfit, lwd = 2, col = "blue")
matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
```

```{r}
table(cut(age, 4))
fit <- lm(wage ~ cut(age, 4), data = Wage)
coef(summary(fit))
```

?