---
title: "ISLR chap6.5."
output: html_notebook
---

# 6.5.1 subset selection

```{r}
library(ISLR2)
#View(Hitters)
names(Hitters)
dim(Hitters)
sum(is.na(Hitters$Salary))
```

disregarding entries with no salary:
```{r}
Hitters2 <- na.omit(Hitters)
dim(Hitters2)
```

subset selection using the `leaps` library:
```{r}
library(leaps)
regfit.full <- regsubsets(Salary ~ ., Hitters2)
summary(regfit.full)
```

let's increase the number of variables (default maxes out at 8 it seems):
```{r}
regfit.full2 <- regsubsets(Salary ~ ., data = Hitters2, nvmax = 19)
reg.summary <- summary(regfit.full2)
reg.summary
```

we can see $R^2$ increase monotonically as more vars are added - which is totally expected:
```{r}
names(reg.summary)
reg.summary$rsq
```

time for some plots!
```{r}
par(mfrow = c(2, 2))
plot(reg.summary$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
which.max(reg.summary$adjr2)
points(11, reg.summary$adjr2[11], col = "red", cex = 2, pch = 20)
```

for the other adjustments (it's pretty cool, they're calculated up front by `leaps`):
```{r}
plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)
points(10, reg.summary$cp[10], col = "red", cex = 2, pch = 20)
which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
points(6, reg.summary$bic[6], col = "red", cex = 2, pch = 20)
```

all 'em plots!
```{r}
plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```
step-wise selection:
```{r}
coef(regfit.full, 6)
regfit.fwd <- regsubsets(Salary ~ ., data = Hitters2, nvmax = 19, method = "forward")
summary(regfit.fwd)
regfit.bwd <- regsubsets(Salary ~ ., data = Hitters2, nvmax = 19, method = "backward")
summary(regfit.bwd)
```


```{r}
coef(regfit.full, 7)
coef(regfit.fwd, 7)
coef(regfit.bwd, 7)
```

let's do this using CV:
```{r}
set.seed(1)
train <- sample(c(TRUE, FALSE), nrow(Hitters2),
replace = TRUE)
test <- (!train)
regfit.best <- regsubsets(Salary ~ ., data = Hitters2[train, ], nvmax = 19)
test.mat <- model.matrix(Salary ~ ., data = Hitters2[test, ])
```

find the number of variables in the best model:
```{r}
val.errors <- rep(NA, 19)
for (i in 1:19) {
coefi <- coef(regfit.best, id = i)
pred <- test.mat[, names(coefi)] %*% coefi
val.errors[i] <- mean((Hitters2$Salary[test] - pred)^2)
}
which.min(val.errors)
coef(regfit.best, which.min(val.errors))
```

making it a fn:
```{r}
predict.regsubsets <- function(object, newdata, id, ...) {
form <- as.formula(object$call[[2]])
mat <- model.matrix(form, newdata)
coefi <- coef(object, id = id)
xvars <- names(coefi)
mat[, xvars] %*% coefi
}
regfit.best <- regsubsets(Salary ~ ., data = Hitters2, nvmax = 19)
coef(regfit.best, 7)
```

using it for $k$-folds
```{r}
k <- 10
n <- nrow(Hitters2)
set.seed(1)
folds <- sample(rep(1:k, length = n))
cv.errors <- matrix(NA, k, 19,
dimnames = list(NULL, paste(1:19)))
```

```{r}
for (j in 1:k) {
best.fit <- regsubsets(Salary ~ .,data = Hitters2[folds != j, ],nvmax = 19)
for (i in 1:19) {
pred <- predict(best.fit, Hitters2[folds == j, ], id = i)
cv.errors[j, i] <- mean((Hitters2$Salary[folds == j] - pred)^2)
}
}
```

```{r}
mean.cv.errors <- apply(cv.errors, 2, mean)
mean.cv.errors
par(mfrow = c(1, 1))
plot(mean.cv.errors, type = "b")
```

10 vars gives the best!
```{r}
reg.best <- regsubsets(Salary ~ ., data = Hitters2, nvmax = 19)
coef(reg.best, 10)
```