---
title: "3.4 with Insurance dataset"
output:
  html_document:
    df_print: paged
---

[Link](https://www.kaggle.com/datasets/mirichoi0218/insurance/) to dataset.

Need to be aware of that a 0 coeff does on categorical variables (it changes the meaning of the intercept)

```{r}
df <- read.csv("insurance.csv")
# clean things up so summary works as expected
df$sex = as.factor(df$sex)
df$smoker = as.factor(df$smoker)
df$region = as.factor(df$region)
summary(df)
```
Data set is roughly equal gender-wise and per region, thought the southeast is a little bit more represented.

Median charges is 9382 but max is 63770 indicating we have a few large ones.

## Q1: Is there a relationship between charges and the factors presented here?

We do this by:
 1. fitting a multiple linear regression model
 2. and testing the null hypothesis (that all coefficients are 0)
 
We can then use the F-statistic and corresponding p-value to decide.

```{r}
lm.fit <- lm(charges ~ ., data=df)
summary(lm.fit)
```

So `F-statistic: 500.8 on 8 and 1329 DF,  p-value: < 2.2e-16` which indicates we can safely reject the null hypothesis.

## Q2: How strong is the relationship?

RSE is 6062! Versus the mean charge of 13270 - that's about 46%. How does that help?

R^2 tells us the predictors explain about 75% of the variance we're seeing.

## Q3: Which predictors are associated with charges?

Interestingly, neither gender nor children nor region seem to have much of an impact.

It'd be worth re-running without the above？

## Q4: How large is the association between each (significant) predictor and charges?

```{r}
confint(lm.fit)
```
Large intervals containing zero can be signs of collinearity. Let's compute the VIF for each var:

```{r}
library(car)
vif(lm.fit)
```
VIF is relatively low, around 1. In a previous passage it was mentioned that anything above 5 would indicate some collinearity.

We can also do separate linear regressions:

```{r}
lm.fitAge <- lm(charges ~ age, data=df)
summary(lm.fitAge)
```

Let's do gender:

```{r}
lm.fitSex <- lm(charges ~ sex, data=df)
summary(lm.fitSex)
```

## Q5: How accurately can we predict future charges?

Here we need a *prediction** interval for an individual response, or a *confidence* interval for an average response.

## Q6: Is the relationhsip linear

Let's plot the residuals:

```{r}
r <- resid(lm.fit)
plot(fitted(lm.fit), r)
```
There are 3 distinct clusters - with one larger one. This indicates some non-linear effect!

For argument's sake, let's do a `pplot` on correlation and we can make a guess:

```{r}
library(GGally)
ggpairs(df)
```

Let's try with a transformation of `bmi`:

```{r}
lm.fitBmiExp <- lm(charges ~ bmi + I(bmi^2) + I(bmi^3), data=df)
summary(lm.fitBmiExp)
```
Meh. Maybe `bmi` isn't such a great indicator of charges?

## Q7: Is there synergy among the predictors?

Let's try an interaction term!

```{r}
lm.fitInter <- lm(charges ~ age + bmi + age*bmi, data=df)
summary(lm.fitInter)
```
No dice either?