---
title: "ISLR Chapter 10"
output:
  html_document:
    df_print: paged
---

# 10.9.1

```{r}
library(ISLR2)
Gitters <- na.omit(Hitters)
n <- nrow(Gitters)
set.seed(13)
ntest <- trunc(n/3)
testid <- sample(1:n, ntest)

lfit <- lm(Salary ~., data=Gitters[-testid,])
lpred <- predict(lfit, Gitters[testid, ])
with(Gitters[testid,],mean(abs(lpred - Salary)))
```

lasso:
```{r}
# -1 to omit the intercept, scale for columns with zero mean and unit variance
x <- scale(model.matrix(Salary ~ . -1, data=Gitters))
y <- Gitters$Salary
library(glmnet)
cvfit <- cv.glmnet(x[-testid,], y[-testid], type.measure = "mae")
cpred <- predict(cvfit, x[testid,], s="lambda.min")
mean(abs(y[testid] - cpred))
```

and now NN!

(the trick to getting this working was to `install_tensorflow()` first, then `install_keras()` - make sure you have done `brew install cmake` first)
```{r}
library(keras)
library(tensorflow)
modnn <- keras_model_sequential() %>%
  layer_dense(units = 50, activation="relu", input_shape=ncol(x)) %>%
  layer_dropout(rate=0.4) %>%
  layer_dense(units=1)
```

a single pipe explanation
```{r}
x <- model.matrix(Salary ~ . -1, data=Gitters) %>% scale()
```

controlling the fitting algo:
```{r}
modnn %>% compile(loss = "mse",
  optimizer = optimizer_rmsprop(),
  metrics = list("mean_absolute_error")
)
```

fitting the model (running multiple times improves the fit)
```{r}
history <- modnn %>% fit(
  x[-testid,], y[-testid], ecpohs=1500, batch_size = 32,
  validation_data = list(x[testid,],y[testid])
)
plot(history)
```

pred:
```{r}
npred <- predict(modnn, x[testid,])
mean(abs(y[testid]-npred))
```

# 10.9.2

load the data
```{r}
mnist <- dataset_mnist()
x_train <- mnist$train$x
g_train <- mnist$train$y
x_test <- mnist$test$x
g_test <- mnist$test$y
dim(x_train)
dim(x_test)
```

one-hot encoding of the class label. note that 784 is 28x28, which is the size (in pixels) of the images
```{r}
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
y_train <- to_categorical(g_train, 10)
y_test <- to_categorical(g_test, 10)
```

scaling by 255 as that's the 8-bit representing grayscale
```{r}
x_train <- x_train / 255
x_test <- x_test / 255
```

NN fit:
```{r}
modelnn <- keras_model_sequential()
modelnn %>%
  layer_dense(units=256, activation="relu", input_shape = c(784)) %>%
  layer_dropout(rate=0.4) %>%
  layer_dense(units = 128, activation="relu") %>%
  layer_dropout(rate=0.3) %>%
  layer_dense(units=10, activation="softmax")
summary(modelnn)
```
fitting algo, + go go go!
```{r}
modelnn %>% compile(loss = "categorical_crossentropy", optimizer = optimizer_rmsprop(), metrics=c("accuracy"))

system.time(
  history <- modelnn %>%
    fit(x_train, y_train, epochs=30, batch_size = 128, validation_split = 0.2)

)
plot(history, smooth=FALSE)
```

accuracy:
```{r}
  accuracy <- function(pred, truth) {
    mean(drop(as.numeric(pred)) == drop(truth))
  }
modelnn %>% predict(x_test) %>% k_argmax() %>% accuracy(g_test)
```

multiclass logistic regression:
```{r}
modellr <- keras_model_sequential() %>%
  layer_dense(input_shape = 784, units=10, activation="softmax")
summary(modellr)
```
fit:
```{r}
modellr %>% compile(loss = "categorical_crossentropy", optimizer=optimizer_rmsprop(), metrics=c("accuracy"))
modellr %>% fit(x_train, y_train, epochs=30, batch_size = 128, validation_split = 0.2)
modellr %>% predict(x_test) %>% k_argmax() %>% accuracy((g_test))
```

# 10.9.3

```{r}
library(keras)
cifar100 <- dataset_cifar100()
names(cifar100)
x_train <- cifar100$train$x
g_train <- cifar100$train$y
x_test <- cifar100$test$x
g_test <- cifar100$test$y
dim(x_train)
range(x_train[1,,, 1])
```

50000 training images, 4 dimensions - 3 colours, each 32x32 8-bit pixels

```{r}
x_train <- x_train/255
x_test <- x_test / 255
y_train <- to_categorical(g_train, 100)
dim(y_train)
```

lookin'
```{r}
library(jpeg)
par(mar=c(0,0,0,0), mfrow=c(5,5))
index <- sample(seq(50000), 25)
for (i in index) plot(as.raster(x_train[i,,,]))
```
now for the model
```{r}
model <- keras_model_sequential() %>%
  layer_conv_2d(filters=32, kernel_size=c(3,3), 
                padding="same",activation="relu",
                input_shape=c(32,32,3)) %>%
  layer_max_pooling_2d(pool_size=c(2,2)) %>%
  layer_conv_2d(filters=64,,kernel_size = c(3,3),
                padding="same",activation="relu") %>%
  layer_max_pooling_2d(pool_size = c(2,2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3),
      padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 256, kernel_size = c(3, 3),
    padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten () %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 100, activation = "softmax")
summary(model)
```
compile!
```{r}
model %>% compile(loss = "categorical_crossentropy",optimizer = optimizer_rmsprop(), metrics = c("accuracy"))
history <- model %>% fit(x_train , y_train , epochs = 30,
batch_size = 128, validation_split = 0.2)
model %>% predict(x_test) %>% k_argmax() %>% accuracy(g_test)
```
# 10.9.4

```{r}
# when this executes in a code block, the cwd isn't the same as your interactive session!
# getwd()
img_dir <- "book_images"
image_names <- list.files(img_dir)
num_images <- length( image_names )
x <- array(dim = c(num_images , 224, 224, 3))
for (i in 1: num_images) {
    img_path <- paste(img_dir , image_names [i], sep = "/")
    img <- image_load(img_path , target_size = c(224 , 224))
    x[i,,, ] <- image_to_array(img)
}
x <- imagenet_preprocess_input(x)
```
the model
```{r}
model <- application_resnet50 ( weights= "imagenet")
summary(model)
```

top-3 preds!
```{r}
pred6 <- model%>% predict(x) %>% imagenet_decode_predictions(top=3)
names(pred6) <- image_names
print(pred6)
```

# 10.10.1

$p=4$, 2 units in the first hidden layer = 4*2 + 4 (biases)
3 units in the second hidden layer = 2*3 + 2 (biases)
1 output

EOF
