---
title: "ISLR Chapter 12"
output:
  html_document:
    df_print: paged
---

# 12.5.1

```{r}
states <- row.names(USArrests)
names(USArrests)
```
examining the data:
```{r}
apply(USArrests, 2, mean)
apply(USArrests, 2, var)
```
PCA
```{r}
pr.out <- prcomp(USArrests, scale=TRUE)
names(pr.out)
pr.out$center
pr.out$scale
```
PCA loadings:
```{r}
pr.out$rotation
```
plot:
```{r}
biplot(pr.out, scale=0)
```

PCAs are only unique up to a sign change - mirror image of the above
```{r}
pr.out$rotation = -pr.out$rotation
pr.out$x = -pr.out$x
biplot(pr.out, scale=0)
```

```{r}
cbind(names(USArrests), pr.out$sdev)
```
variance:
```{r}
pr.var <- pr.out$sdev^2
pr.var
```

proportion:
```{r}
pve <- pr.var/sum(pr.var)
pve
```

plot PVEs:
```{r}
par(mfrow=c(1,2))
plot(pve , xlab = " Principal Component",
ylab = "Proportion of Variance Explained", ylim = c(0, 1),
type = "b")
plot(cumsum(pve), xlab = "Principal Component",
ylab = "Cumulative Proportion of Variance Explained",
ylim = c(0, 1), type = "b")
```
# 12.5.2 Matrix Completion

```{r}
X <- data.matrix(scale(USArrests))
pcob <- prcomp(X)
summary(pcob)
```
PC1 explains 62% of the variance. we now use singular value decomposition to solve 12.6
```{r}
sX <- svd(X)
names(sX)
round(sX$v, 3)
```
this is equal to the loading matrix up to the sign flip:
```{r}
pcob$rotation
```
```{r}
t(sX$d * t(sX$u))
pcob$x
```

omit one of the 4 features for 20 entries:
```{r}
nomit <- 20
set.seed(15)
# states selected to have missing values
ina <- sample(seq(50), nomit)
# which of the 4 features to have missing value
inb <- sample(1:4, nomit, replace=TRUE)
Xna <- X
index.na <- cbind(ina, inb)
Xna[index.na] <- NA
```

step 2 of algo 12.1:
```{r}
fit.svd <- function (X, M = 1) {
  svdob <- svd(X)
  with(svdob,
       u[, 1:M, drop = FALSE] %*%
      (d[1:M] * t(v[, 1:M, drop = FALSE ]))
  )
}
```

replace missing entries with column means:
```{r}
Xhat <- Xna
xbar <- colMeans(Xna, na.rm=TRUE)
Xhat[index.na] <- xbar[inb]
```

measure the progress of each iteration
```{r}
thresh <- 1e-7
rel_err <- 1
iter <- 0
ismiss <- is.na(Xna)
mssold <- mean (( scale(Xna , xbar , FALSE)[! ismiss ]) ^2)
mss0 <- mean(Xna [! ismiss ]^2)
```

relative error will be `(mssold - mss) / mss0`

# 12.5.3 Clustering

k-means

```{r}
set.seed(2)
x <- matrix(rnorm(50*2), ncol=2)
x[1:25,1] <- x[1:25, 1] + 3
x[1:25,1] <- x[1:25, 1] - 4
```
EOF

# 12.5.4 NCI60

```{r}
library(ISLR2)
```
